---
title: "R Notebook"
output: html_notebook
---

Installing packages that will be useful for our mapping process:
```{r}
install.packages("tidyverse",repos = 'https://CRAN.R-project.org/package=tidyverse')
install.packages("maptools")
install.packages(c("OpenStreetMap", "tmap"))
install.packages("classInt")
install.packages("readr")

install.packages("munsell")


library(tmaptools)
library(RColorBrewer)
library(sp)
library(classInt)
library(rgeos)
library(tmap)
library(tmaptools)
library(sf)
library(rgdal)
library(readr)
library(tidyverse)


```



PART 1 / London boroughs boundaries

Reading shp file of London Boundaries. Renames the column with the area codes into "code"

```{r}
eng_bound <- read_shape("C:/Users/User/Desktop/work/casa/0005 GIS/1. week 4 assignment/London map/eng_bound.shp", as.sf = TRUE)

names(eng_bound)[1]<-"code"

qtm(eng_bound)

```



PART 2 / Enterprise births data

Reads csv with the enterprises births data and cleans characters from numeric values. Checks if the data is read correctly. Renames column with area code. From the dataframe keeps only the rows that include the London Borough Codes (starting with E09)


```{r}

business_original<- read_csv("C://Users/User/Desktop/work/casa/0005 GIS/1. week 4 assignment/startups/edited_birth_enterprises.csv", na = "n/a")

datatypelist_enterpr <- data.frame(cbind(lapply(business_original,class)))

names(business_original)[1] <- c("code1")

London_data <- business_original[grep("^E09",business_original$code1),]


```


PART 3

Joins map of London and enterprise data according to their common area code field. Plots a sample map with the mean of enterprises registered throughout the years.


```{r}
class(London_data)

enterprises_map <- append_data(eng_bound,London_data,key.shp="code",key.data="code1",ignore.duplicates = TRUE)

class(enterprises_map)

```


PART 4 

Starting plotting the maps. Creating a sample map of London boroughs, trying different linetypes.

```{r}
map1 <-  tm_shape(enterprises_map) + tm_borders(lty = 2)
print(map1)

```



Rearranges the dataframe of business_original so that every column (of years between 2004-2016) consists a single one, compiled above the previous one. From the final dataframe, it keeps the geometry field and only the rows which contain London Borough codes

```{r}

library(reshape2)
library(dplyr)
london_melt <- melt(business_original,id.vars = 1:2, measure.vars = c(3,7,11,15))



london_melt <- left_join(london_melt,eng_bound,by = c("code1" = "code"))

london_melt <- london_melt[,c(1:4,10)]


library(sf)

london_melt <- london_melt[grep("^E09",london_melt$code1),]
london_melt <- st_as_sf(london_melt)




```


Trying a sample faceted map with tmap 

```{r}
library(tmap)
sample_map=tm_shape(london_melt) +
    tm_polygons("value", palette = "-magma",n=8) +
    tm_facets(by = "variable")+tm_layout(panel.labels = c('2004','2008','2012','2016'), panel.label.color = 'white',panel.label.size = 1,panel.label.bg.color = '#79839c',legend.show = TRUE, legend.text.size = 0.5,title="My Title", title.size=1)+tm_compass(type="arrow",size=1.5,fontsize = 0.8)

print(sample_map)
```


PART 5

Creating the final faceted plot using ggplot package

```{r}

install.packages("ggplot2")
library("ggplot2")
install.packages("gridExtra")
library(gridExtra)
library(grid)
library(ggplot2)

library(ggsn)
install.packages("viridis")
library(viridis)



theme1 <- theme(strip.text.x= element_text(size=11, angle=0,colour="#262321"),strip.background = element_rect(colour="#c3c3c2", fill="#7383a5"),legend.background = element_rect(fill="white",size=0.7, linetype="solid",colour ="#2f446f"),legend.position=c("right"),legend.box.margin=margin(c(70,70,70,30)),panel.background = element_rect(fill = '#e7e6f8', colour = '#f2f2f2'),panel.grid.major = element_line(colour = "#ead595"),axis.text.x = element_text(size=9,colour="#d8680f"),axis.text.y = element_text(size=9,colour="#d8680f"),title =element_text(size=14),panel.spacing = unit(1, "lines"))

labels <- c("2004" = "2004", "2008" = "2008","2012" = "2012","2016" = "2016")


a <- ggplot()+geom_sf(mapping = aes(geometry=geometry,fill=value),data = london_melt)+facet_wrap(~ variable, ncol = 2,labeller=labeller(variable = labels))+theme1+ggtitle("New enterprises in London, by Borough")+labs(caption = "Source: Office for National Statistics")+scale_fill_viridis(option = "A",direction=-1)+ labs(fill = "Number of enterprise births")

north2(a, x = 0.76, y = 0.3, scale = 0.1, symbol = 1)
print(a)
pdf(file="enterpriseBirths.pdf",width=10.5,height=8)


```




